Generalized Bradley-Terry models

### 2025-03-13 inst/jss_hyper3_real.Rnw
minor tweaks

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Robin K. S. Hankin\\
  University of Stirling\\
  E-mail: \email{hankin.robin@gmail.com}\\
  URL: \url{https://academics.aut.ac.nz/robin.hankin}
}

@@

The Bradley-Terry model for datasets involving paired comparisons has
wide uptake in the {R} community.  The canonical problem is to
consider~$n$ players who compete against one another; the basic
inference problem is to estimate
numbers~$\mathbf{p}=\left({p_1,\ldots,p_n}\right)$, $p_i\geqslant 0$,
$\sum p_i=1$ so that
$\Prob(\mbox{$i$~beats~$j$})=\frac{p_i}{p_i+p_j}$.  Information about
the~$p_i$ may be obtained from the results of paired comparisons
between the players.

@@

where $w_{12}$ and $w_{34}$ are the number of wins by each team
respectively.  Then $\lambda$ and $\mu$ represent the effectiveness of
$p_1$'s special move and the enhancement afforded to $p_4$ by the
drug, with values $>1$ corresponding to a positive effect.  In this
paper I generalize the {\tt hyper2} suite for Bradley-Terry
likelihoods~\citep{hankin2017_rnw} to accommodate this class of
multiplicative factors.  The most general likelihood function would be
proportional to

@@

Further, we may calculate a support function over the
$(\lambda,D)$-plane, shown in figure~\ref{contourlambdaD}.  For
example, we may test the hypothesis that $\lambda=1$ [corresponding to
  the right hand side of figure~\ref{contourlambdaD}] for which the
support is about \Sexpr{round(f(maxv) - jj$objective,2)}, far
exceeding the two units of support per degree of freedom
criteron~\citep{edwards1992}.  Alternatively, we can observe that this
lies in the tail region of its asymptotic null $\chi^2_4$ distribution
with a $p$-value of about \Sexpr{signif(pchisq(f(maxv) -
  jj$objective,lower.tail=FALSE,df=2),digits=2)}.

@@

As an example, consider figure~\ref{tikzabcde}.  This shows a partial
probability tree diagram for some of the $5!=120$ possible order
statistics.  The standard Plackett-Luce likelihoods have been modified
to account for two groups of mutually supporting runners:
$\left\lbrace a,b,c\right\rbrace$ with support term $\lambda$, and
$\left\lbrace d,e\right\rbrace$ with support term $\mu$.  From START,
the first runner to cross the finishing line has standard
Plackett-Luce probability.  Taking the top path as an example, we see
that the likelihood function for $a\succ b\succ c\succ d\succ e$ would
be

>>>>

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Robin K. S. Hankin\\
  University of Stirling\\
  E-mail: \email{hankin.robin@gmail.com}
}

@@

The Bradley-Terry model for datasets involving paired comparisons has
wide uptake in the {R} community.  The canonical case is to
consider~$n$ players who compete against one another in pairs; the
basic inference problem is to estimate
numbers~$\mathbf{p}=\left({p_1,\ldots,p_n}\right)$, $p_i\geqslant 0$,
$\sum p_i=1$ so that
$\Prob(\mbox{$i$~beats~$j$})=\frac{p_i}{p_i+p_j}$.  Information about
the~$p_i$ may be obtained from the results of paired comparisons
between the players.

@@

where $w_{12}$ and $w_{34}$ are the number of wins by each team
respectively.  Then $\lambda$ would represent the effectiveness of
$p_1$'s special move, and $\mu$ the enhancement afforded to $p_4$ by
the drug; values $>1$ would correspond to a positive effect.  In this
paper I generalize the {\tt hyper2} suite for Bradley-Terry
likelihoods~\citep{hankin2017_rnw} to accommodate this class of
multiplicative factors.  The most general likelihood function would be
proportional to

@@

Further, we may calculate a support function over the
$(\lambda,D)$-plane, shown in figure~\ref{contourlambdaD}.  For
example, we may test the hypothesis that $\lambda=1$ [corresponding to
  the right hand side of figure~\ref{contourlambdaD}] for which the
support is about \Sexpr{round(f(maxv) - jj$objective,2)}, far
exceeding the two units of support per degree of freedom
criteron~\citep{edwards1992}.  Alternatively, we can observe that this
lies in the tail region of its asymptotic null $\chi^2_2$ distribution
with a $p$-value of about \Sexpr{signif(pchisq(f(maxv) -
  jj$objective,lower.tail=FALSE,df=2),digits=2)}.

@@

As an example, consider figure~\ref{tikzabcde}.  This shows a partial
probability tree diagram for some of the $5!=120$ possible order
statistics for five competitors.  The standard Plackett-Luce
likelihoods have been modified to account for two groups of mutually
supporting runners: $\left\lbrace a,b,c\right\rbrace$ with support
term $\lambda$, and $\left\lbrace d,e\right\rbrace$ with support term
$\mu$.  From START, the first runner to cross the finishing line has
standard Plackett-Luce probability.  Taking the top path as an
example, we see that the likelihood function for $a\succ b\succ c\succ
d\succ e$ would be
